---
title: "Thermal niche model reported in Gutaker et al. 2020"
author:
  - R. Kyle Bocinsky
  - Jade d'Alpoim Guedes
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
   cores: 2
   clean: FALSE
   google_maps_elevation_api_key: !r Sys.getenv('google_maps_elevation_api_key')
output: 
    bookdown::html_document2:
      code_folding: hide
      df_print: paged
      fig_caption: yes
      toc: true
      toc_float: true
      toc_depth: 3
editor_options: 
  chunk_output_type: console
---

This RMarkdown document creates the thermal niche model reported in 

> Gutaker et al. *In review.*

```{r, setup, echo = FALSE, cache = FALSE}
# params <- list(cores = 2,
#                clean = FALSE)

# Set the knitting behavior
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      collapse = TRUE, 
                      # cache = FALSE,
                      cache = !params$clean,
                      cache.lazy = FALSE,
                      results = 'hold',
                      out.width = "100%",
                      fig.height = 5,  
                      comment = "#>",
                      fig.path = here::here("analysis/figures/")
)

library(gutaker2020)
library(sf)
library(magrittr)
library(fields)

# Set the behavior for printing scientific notation
# options(digits = 3, scipen = -2)
options(scipen = 999,
        knitr.table.format = "html")

# Force Raster to load large rasters into memory
raster::rasterOptions(chunksize = 2e+08,
                      maxmemory = 2e+09)

# Create the plan for (possibly parallel) execution
future::plan(future::multisession, 
             workers = min(future::availableCores(), 
                           params$cores),
             gc = TRUE)

raw_data <- here::here("analysis/data/raw_data/")
derived_data <- here::here("analysis/data/derived_data/")
figures <- here::here("analysis/figures/")

# Create output directories
dir.create(raw_data,
           recursive = TRUE,
           showWarnings = FALSE)
dir.create(derived_data,
           recursive = TRUE,
           showWarnings = FALSE)
dir.create(figures,
           recursive = TRUE,
           showWarnings = FALSE)

```

## Set parameters {-}
In this section we define the study area by a bounding box, the calendar years used to define climatology (here, 1961--1990), and a plotting theme for **ggplot**.

```{r set-params}
# Set a series of colors for the crops
crop_colors <- RColorBrewer::brewer.pal(6,"Dark2")

# A function to set an objects class
set_class <- function(x, classes){
  class(x) <- classes
  return(x)
}

#Define the study region
ASIA_poly <-
  raster::extent(55,150,-11,60) %>%
  FedData::polygon_from_extent("+proj=longlat") %>%
  sf::st_as_sf()

# Set the calibration period for paleoclimate reconstructions
calibration.years <- 1961:1990

# A ggplot2 theme for Nature Publishing Group
nature_theme <- ggplot2::theme(panel.grid.major = ggplot2::element_line(size = 0.5, 
                                                                        color = "grey"),
                               axis.line = ggplot2::element_line(size = 0.7, 
                                                                 color = "black"),
                               legend.position = c(0.85, 0.7),
                               text = ggplot2::element_text(size = 14))
```

## Load ETOPO5 grids {-}
Indicator kriging takes place on the ETOPO5 digital elevation model. Here, we load the ETOPO5 data from the **geomapdata** package, then mask out the oceans adn large lakes, as well as islands outside of mainland Asia.
```{r etopo5}
message("Preparing the ETOPO5 grid-aligned dataset")

time_check <-  Sys.time()

ne_dir <- paste0(raw_data,"NaturalEarth/")

if(!params$clean & file.exists(paste0(raw_data,"ASIA_rast_etopo5.tif"))){
  
  ASIA_rast_etopo5 <- raster::raster(paste0(raw_data,"ASIA_rast_etopo5.tif"))
  
}else{
  
  # Get the Natural Earth country lakes data
  dir.create(ne_dir, 
             showWarnings = FALSE, 
             recursive = TRUE)
  
  FedData::download_data(
    url = "http://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/physical/ne_10m_lakes.zip",
    destdir = ne_dir
  )
  
  unzip(
    paste0(ne_dir,"ne_10m_lakes.zip"), 
    exdir = paste0(ne_dir,"ne_10m_lakes/")
  )
  
  ne_10m_lakes <-  paste0(ne_dir,"ne_10m_lakes/ne_10m_lakes.shp") %>%
    sf::st_read() %>%
    sf::st_transform(sf::st_crs(ASIA_poly)) %>%
    sf::st_intersection(ASIA_poly)
  
  # Get the Natural Earth country boundaries data
  dir.create(
    paste0(ne_dir,"ne_10m_admin_0_countries_lakes/"),
    showWarnings = FALSE, 
    recursive = TRUE
  )
  FedData::download_data(
    url = "http://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/cultural/ne_10m_admin_0_countries_lakes.zip",
    destdir = ne_dir
  )
  
  unzip(
    paste0(ne_dir,"ne_10m_admin_0_countries_lakes.zip"), 
    exdir = paste0(ne_dir,"ne_10m_admin_0_countries_lakes/")
  )
  
  ne_10m_admin_0_countries_lakes <- 
    paste0(ne_dir,"ne_10m_admin_0_countries_lakes/ne_10m_admin_0_countries_lakes.shp") %>%
    sf::st_read() %>%
    sf::st_transform(sf::st_crs(ASIA_poly)) %>%
    sf::st_intersection(ASIA_poly) %>%
    dplyr::filter(!(NAME %in% c("Scarborough Reef",
                                "Maldives",
                                "Spratly Is.",
                                "Oman",
                                "United Arab Emirates",
                                "Saudi Arabia")))
  
  library(geomapdata)
  data("ETOPO5")
  
  ASIA_rast_etopo5 <- ETOPO5 %>%
    t() %>%
    raster::raster(xmn = 0,
                   xmx = 360,
                   ymn = -90,
                   ymx = 90,
                   crs = sp::CRS("+proj=longlat +ellps=clrk66 +no_defs")) %>%
    raster::crop(ASIA_poly %>%
                   sf::st_transform("+proj=longlat +ellps=clrk66 +no_defs") %>%
                   as("Spatial")) %>%
    raster::mask(ne_10m_admin_0_countries_lakes %>%
                   sf::st_transform("+proj=longlat +ellps=clrk66 +no_defs") %>%
                   as("Spatial")) %>%
    raster::mask(ne_10m_lakes %>%
                   sf::st_transform("+proj=longlat +ellps=clrk66 +no_defs") %>%
                   as("Spatial"),
                 inverse = TRUE) %T>%
    raster::writeRaster(filename = paste0(raw_data,"ASIA_rast_etopo5.tif"),
                        datatype="INT2S",
                        options=c("COMPRESS=DEFLATE", "ZLEVEL=9", "INTERLEAVE=BAND", "PHOTOMETRIC=MINISWHITE"),
                        overwrite=T,
                        setStatistics=FALSE)
  rm(ETOPO5)
}

countries <- 
  paste0(ne_dir,"ne_10m_admin_0_countries_lakes/ne_10m_admin_0_countries_lakes.shp") %>%
  sf::st_read()  %>%
  sf::st_transform(sf::st_crs(ASIA_poly)) %>%
  sf::st_intersection(ASIA_poly) %>%
  dplyr::filter(!(NAME %in% c("Scarborough Reef",
                              "Maldives",
                              "Spratly Is.",
                              "Oman",
                              "United Arab Emirates",
                              "Saudi Arabia"))) %>%
  dplyr::select(NAME, geometry)

message("ETOPO5 grid-aligned dataset preparation complete: ", capture.output(Sys.time() - time_check))

```

## Prepare the GHCN data {-}
Here, we download daily weather records from the Global Historical Climatology Network, 
```{r ghcn}
## Downloads and cleans daily climate records from the Global Historical Climate Database.
message("Preparing the daily climate records from the Global Historical Climate Database")
time_check <-  Sys.time()
GHCN.data.final <- prepare_ghcn(region = ASIA_poly %>%
                                  as("Spatial"),
                                label = "ASIA_poly",
                                calibration.years = calibration.years,
                                google_maps_elevation_api_key = params$google_maps_elevation_api_key,
                                raw_dir = paste0(raw_data,"ghcn/"),
                                derived_dir = paste0(derived_data,"ghcn/"),
                                force.redo = params$clean)
message("GHCN preparation complete: ", capture.output(Sys.time() - time_check))
## An example of plotting the GHCN data
# climate_plotter(data = GHCN.data.final, station = "CHM00051334", element = "TMIN")

```

## Prepare the Marcott data {-}
```{r marcott}
# Run the script that transforms the Marcott et al. 2013 data into standard scores.
message("Preparing the Marcott et al. 2013 data")
time_check <-  Sys.time()
marcott2013 <- prepare_marcott(calibration.years = calibration.years,
                               raw_dir = paste0(raw_data,"marcott2013/"),
                               derived_dir = paste0(derived_data,"marcott2013/"))
message("Marcott et al. 2013 data preparation complete: ", capture.output(Sys.time() - time_check))

```

## Modulating climatology by SD {-}
```{r sd-modulation}
#### Calculating growing degree days ####
message("Modulating local climatology by Marcott et al. 2013 data")
time_check <-  Sys.time()
# How often to sample GDD, in z-space, for model tuning
# Here, we sample from -20 to 20 SD, at 1 SD interval
sample.points <- -20:20

# Read in data on different crop GDD needs
crop_GDD <- 
  readr::read_csv(here::here("inst/crops.csv"),
                  col_types = readr::cols(
                    cultivar_long = readr::col_character(),
                    cultivar = readr::col_character(),
                    crop_long = readr::col_character(),
                    crop = readr::col_character(),
                    t_base = readr::col_double(),
                    min_gdd = readr::col_integer()
                  )) %T>%
  readr::write_csv(paste0(derived_data,"/crops.csv"))

# Transform GHCN data to GDDs of each base, and modulate to Marcott
GDDs <- sort(unique(crop_GDD$t_base))

GHCN.GDD.incremented.sd <-
  furrr::future_map(GDDs,
                    # purrr::map(GDDs,
                    .options = furrr::future_options(
                      packages = c("magrittr",
                                   "gutaker2020"),
                      globals = c("sample.points",
                                  "GHCN.data.final")),
                    .f = function(base){
                      purrr::map_dfr(sample.points,
                                     function(change){
                                       
                                       GHCN.GDDs <-
                                         purrr::map_dbl(GHCN.data.final$climatology,
                                                        function(station){
                                                          return(sdModulator(data.df = station,
                                                                             temp.change.sd = change,
                                                                             t.base = base,
                                                                             t.cap = 30))
                                                        })
                                       
                                       
                                       names(GHCN.GDDs) <- names(GHCN.data.final$climatology)
                                       
                                       return(tibble::tibble(SD_change = change,
                                                             ID = names(GHCN.GDDs),
                                                             GDD = GHCN.GDDs))
                                       
                                     }) %>%
                        dplyr::left_join(GHCN.data.final$spatial %>%
                                           tibble::as_tibble() %>%
                                           magrittr::set_names(c("ID",
                                                                 "NAME",
                                                                 "elevation",
                                                                 "x",
                                                                 "y")),
                                         by = "ID")
                      
                    })

names(GHCN.GDD.incremented.sd) <- GDDs

# stop the cluster (will free memory)
# stopCluster(cl)
message("Modulation of local climatology by Marcott et al. 2013 data complete: ", capture.output(Sys.time() - time_check))

```

## Generate the crop niche models {-}
```{r generate-niche-models}
message("Calculating indicator Krige models")
time_check <-  Sys.time()
# Create a spatialPointsDataFrame of the etopo5 data, and convert to WGS84 ellipsoid
ASIA_rast_etopo5.sp <-
  ASIA_rast_etopo5 %>%
  magrittr::set_names("elevation") %>%
  raster::rasterToPoints(spatial = T) %>%
  sp::spTransform(sp::CRS(raster::projection(GHCN.data.final$spatial)))

# A function that generates the kriging model, then predicts
krige_and_predict <- function(dt){
  model <- fields::mKrig(x = dt[,c("x","y")],
                         y = dt$GDD_thresh,
                         Z = dt$elevation,
                         Covariance = "Exponential",
                         Distance = "rdist.earth")
  
  prediction <- ASIA_rast_etopo5.sp %>%
    tibble::as_tibble() %>%
    dplyr::mutate(chunk = rep(1:ceiling(nrow(.)/chunk_size),length.out = nrow(.)) %>%
                    sort()
    ) %>%
    dplyr::group_by(chunk) %>%
    dplyr::do(prediction = fields::predict.mKrig(model,
                                                 xnew = .[,c("x","y")],
                                                 Z = .$elevation) %>%
                as.vector()) %$%
    prediction %>%
    unlist()
  
  return(list(model = model, prediction = prediction))
  
}

# Calculate gdd kriging models for each crop
gdd_model_files <- paste0(derived_data,"models/",crop_GDD$cultivar,"_models.rds")

if(params$clean){
  unlink(paste0(derived_data,"models/"), recursive = TRUE, force = TRUE)
  dir.create(paste0(derived_data,"models/"), recursive = TRUE, showWarnings = F)
  crop_GDD_run <- crop_GDD
}else{
  dir.create(paste0(derived_data,"models/"), recursive = TRUE, showWarnings = F)
  crop_GDD_run <- crop_GDD[!file.exists(gdd_model_files),]
}

if(nrow(crop_GDD_run) == 0){
  message("All indicator Krige models have already been calculated. Continuing.")
}else message("Calculating indicator Krige models for ",
              nrow(crop_GDD_run),
              " cultivars:\n",
              paste0(capture.output(crop_GDD_run), collapse = "\n"))

# A function to reduce the size of a loess model by half
skinny.loess <- function(x){
  x[c("fitted",
      "residuals",
      "enp",
      "one.delta",
      "two.delta",
      "trace.hat",
      "call",
      "terms",
      "xnames")] <- NULL
  return(x)
}

# A function of correct the indication predictions and estimate a smooth
# monotonic function
# This first uses isotonic regression, then loess smoothing with a degree of 1
smooth.preds <- function(y){
  y[y<0] <- 0
  y[y>1] <- 1
  y <- loess(isoreg(y~sample.points)$yf~sample.points, span=0.1, degree=1) %>%
    skinny.loess()
  return(y)
}

if(nrow(crop_GDD_run) > 0){
  options(dplyr.show_progress = FALSE)
  chunk_size <- 10000
  
  gdd.models <-
    furrr::future_map(1:nrow(crop_GDD_run),
                      .options = furrr::future_options(
                        packages = c("fields",
                                     "dplyr",
                                     "magrittr",
                                     "readr",
                                     "gutaker2020")),
                      .f = function(crop){
                        
                        # Threshold for indicator kriging
                        GHCN.GDD.incremented.sd[[as.character(crop_GDD_run[crop,"t_base"])]] %>%
                          dplyr::mutate(GDD_thresh = {GDD >= as.numeric(crop_GDD_run[crop,"min_gdd"])}) %>%
                          dplyr::group_by(SD_change) %>%
                          dplyr::do(out_preds = krige_and_predict(.)) %$%
                          out_preds %>%
                          sapply("[[","prediction") %>%
                          apply(1,smooth.preds) %>%
                          tibble::tibble(model = .) %>%
                          maptools::spCbind(ASIA_rast_etopo5.sp,.) %>%
                          readr::write_rds(paste0(derived_data,"models/",
                                                  crop_GDD_run[crop,"cultivar"],
                                                  "_models.rds"), 
                                           compress = "xz")
                        
                        return(paste0(derived_data,"models/",
                                      crop_GDD_run[crop,"cultivar"],
                                      "_models.rds"))
                        
                      })
  
}

rm(ASIA_rast_etopo5,
   ASIA_rast_etopo5.sp)

message("Calculation indicator Krige models complete: ", capture.output(Sys.time() - time_check))

```

## Predict crop niches through time {-}
```{r predict-niches}
## Predicting crop niche from smoothed Krige models
# Calculate niches for each crop using the Marcott et al. 2013.
# create the cluster for parallel computation
message("Generating niche reconstructions")
time_check <-  Sys.time()
if(params$clean){
  unlink(paste0(derived_data,"recons/"), recursive = TRUE, force = TRUE)
}
dir.create(paste0(derived_data,"recons/"), recursive = TRUE, showWarnings = F)

# Create the plan for (possibly parallel) execution
future::plan(future::multisession, 
             workers = max(
               floor(
                 min(
                   future::availableCores(), 
                   params$cores)/2), 
               1),
             gc = TRUE)

furrr::future_walk(crop_GDD$cultivar,    
                   .options = furrr::future_options(
                     packages = c("fields",
                                  "dplyr",
                                  "magrittr",
                                  "sp",
                                  "readr",
                                  "gutaker2020")),
                   function(crop){
                     
                     if(file.exists(paste0(derived_data,"recons/",
                                           crop,
                                           "_recons.rds")))
                       return()
                     
                     if(!file.exists(paste0(derived_data,"models/",
                                            crop,
                                            "_models.rds"))) 
                       stop("Models for ",
                            crop,
                            " are missing! Aborting.")
                     
                     Zs <- c("Z_Lower","Z","Z_Upper")
                     
                     crop.models <- readr::read_rds(paste0(derived_data,"models/",
                                                           crop,
                                                           "_models.rds"))
                     
                     purrr::walk(Zs, 
                                 function(z){
                                   suppressWarnings(crop.models@data %$%
                                                      model %>%
                                                      purrr::map(
                                                        function(x){
                                                          x %>%
                                                            predict(newdata = marcott2013[[z]]) %>%
                                                            magrittr::multiply_by(100) %>%
                                                            round() %>%
                                                            as.integer()
                                                        }) %>%
                                                      do.call(rbind, .) %>%
                                                      tibble::as_tibble() %>%
                                                      magrittr::set_colnames(marcott2013$YearBP) %>%
                                                      new("SpatialPointsDataFrame",
                                                          data = .,
                                                          coords.nrs = crop.models@coords.nrs,
                                                          coords = crop.models@coords,
                                                          bbox = crop.models@bbox,
                                                          proj4string = crop.models@proj4string) %>%
                                                      as("SpatialPixelsDataFrame") %>%
                                                      raster::brick() %>%
                                                      raster::setZ(marcott2013$YearBP,
                                                                   name="Years BP"))
                                 }) %>%
                       magrittr::set_names(Zs) %>%
                       readr::write_rds(paste0(derived_data,"recons/",
                                               crop,
                                               "_recons.rds"),
                                        compress = "xz",
                                        compression = 9)
                     
                     return()
                   })

message("Generation of niche reconstructions complete: ", capture.output(Sys.time() - time_check))

```

## Combine similar crop niches {-}
```{r combine-niches}
## Combining crop niches from similar crops by taking an arithmatic mean
message("Combining like crop niches")
time_check <-  Sys.time()

# Force Raster to load large rasters into memory
raster::rasterOptions(chunksize = 2e+10,
                      maxmemory = 2e+11)

crop_GDD %>%
  split(as.factor(crop_GDD$crop)) %>%
  purrr::walk(function(x){
    
    if(file.exists(paste0(derived_data,"recons/all_",x$crop[[1]],".rds"))) return()
    n_crops <- length(x$crop)
    
    if(n_crops == 1){
      file.copy(paste0(derived_data,"recons/",
                       x$cultivar[[1]],
                       "_recons.rds"),
                paste0(derived_data,"recons/all_",
                       x$crop[[1]],
                       ".rds"))
      return()
    }
    
    x %$%
      cultivar %>%
      purrr::map(function(cultivar){
        paste0(derived_data,"recons/",cultivar,"_recons.rds") %>%
          readr::read_rds()
      }) %>%
      purrr::transpose() %>%
      purrr::map(function(x){
        x %>%
          Reduce(f = "+", x = .) %>%
          magrittr::divide_by(n_crops) %>%
          round() %>%
          raster::readAll() %>%
          magrittr::set_names(names(x[[1]]))
      }) %>%
      readr::write_rds(paste0(derived_data,"recons/all_",
                              x$crop[[1]],
                              ".rds"),
                       compress = "xz",
                       compression = 9)
  })

message("Combining like crop niches complete: ", capture.output(Sys.time() - time_check))

```



```{r plot-niches-through-time, cache = FALSE}
## Plotting cultivar niche
# create the cluster for parallel computation
message("Plotting cultivar niche reconstructions")
time_check <-  Sys.time()

dir.create(paste0(figures,"cultivar_niches/"),
           recursive = TRUE,
           showWarnings = FALSE)

breaks <- seq(0, 100, 10)

pal <- viridisLite::viridis(length(breaks) - 1)

gdd.recons <- 
  purrr::map_chr(1:nrow(crop_GDD),
                 function(n){
                   
                   cultivar <- crop_GDD[n,]$cultivar
                   
                   title <- stringr::str_c(crop_GDD[n,]$cultivar_long,
                                           " \u2014 Required GDD: ",
                                           crop_GDD[n,]$min_gdd,
                                           " at a ",
                                           crop_GDD[n,]$t_base,
                                           "°C base")
                   
                   rasts <- readr::read_rds(paste0(derived_data,"recons/",cultivar,"_recons.rds")) %>%
                     purrr::map(function(x){
                       x %>%
                         magrittr::extract2(which(.@z$`Years BP` > 1000)) %>%
                         # raster:::readAll() %>%
                         magrittr::extract2(raster::nlayers(.):1) 
                     })
                   
                   
                   years <- rasts[[1]] %>%
                     names() %>%
                     gsub(pattern = "X",
                          replacement = "",
                          x = .) %>%
                     as.numeric()
                   
                   if(!file.exists(paste0(figures,"cultivar_niches/",cultivar,".pdf")))
                     gutaker2020:::space_time_plot(
                       the_brick = rasts$Z,
                       the_brick_lower = rasts$Z_Lower,
                       the_brick_upper = rasts$Z_Upper,
                       out_file = paste0(figures,"cultivar_niches/",cultivar,".pdf"),
                       title = title,
                       time = years,
                       timelim = c(max(years),min(years)),
                       timeaxis =  seq(from = max(years)-500,
                                       to = min(years),
                                       by = -500),
                       timelab = "Years BP",
                       zbreaks = breaks,
                       zlab = "Probability of being in niche",
                       zaxis = seq(0,100,10),
                       zcolors = pal
                     )
                   
                   if(!file.exists(paste0(figures,"cultivar_niches/",cultivar,".mp4")))
                     gutaker2020:::space_time_video(                       
                       the_brick = rasts$Z,
                       the_brick_lower = rasts$Z_Lower,
                       the_brick_upper = rasts$Z_Upper,
                       out_file = paste0(figures,"cultivar_niches/",cultivar,".mp4"),
                       title = title,
                       time = years,
                       timelim = c(max(years),min(years)),
                       timeaxis =  seq(from = max(years)-500,
                                       to = min(years),
                                       by = -500),
                       timelab = "Years BP",
                       zbreaks = breaks,
                       zlab = "Probability of being in niche",
                       zaxis = seq(0,100,10),
                       zcolors = pal
                     )
                   
                   return(paste0(figures,"cultivar_niches/",cultivar,".pdf"))
                   
                 })

message("Plotting of cultivar niche reconstructions complete: ", capture.output(Sys.time() - time_check))

```

## Generate niche videos {-}
```{r niche-videos}
# Create a static plot for paper publication
# A function to extract data from a crop raster brick
tidyCrop <- function(x, years){
  x %>%
    paste0(derived_data,"recons/all_",.,".rds") %>%
    readr::read_rds() %$%
    Z %>%
    magrittr::extract2(stringr::str_c("X",years)) %>%
    as("SpatialPixelsDataFrame") %>%
    tibble::as_tibble() %>%
    tidyr::gather(Year, Niche, -x:-y) %>%
    dplyr::rename(Longitude = x,
                  Latitude = y) %>%
    dplyr::mutate(Year = stringr::str_remove(Year, "X") %>%
                    factor(levels = years) %>%
                    forcats::fct_relabel(function(x){stringr::str_c(x," cal. BP")}),
                  Crop = x)
}

breaks <- seq(0, 100, 10)

pal <- (RColorBrewer::brewer.pal(10, "Spectral") %>%
          rev() %>%
          colorRampPalette(.))(length(breaks) - 1)

# A function to create a plot of years and crops
facet_niche <- function(crops, years){
  
  these_rasters <- crops %>%
    purrr::map(tidyCrop, years = years) %>%
    dplyr::bind_rows() %>%
    dplyr::mutate(Crop = dplyr::recode_factor(Crop,
                                              rice = "Rice"),
                  Niche = Niche  %>%
                    cut(breaks = breaks,
                        include.lowest = TRUE)) 
  
  p <- these_rasters %>%
    ggplot2::ggplot() +
    ggplot2::geom_raster(mapping = ggplot2::aes(x = Longitude,
                                                y = Latitude,
                                                fill = Niche),
                         show.legend = TRUE) +
    ggplot2::coord_quickmap() +
    ggplot2::facet_grid(Crop ~ Year,
                        switch = "y")
  
  return(p)
}

message("Plotting of crop niche reconstructions complete: ", capture.output(Sys.time() - time_check))

```

## Are GDD differences between Japan and Thailand stable across the reconstruction? {-}
```{r niche-compare}
japan_thai <- countries %>%
  dplyr::filter(NAME %in% c("Japan","Thailand"))

# # Get the pixel counts
# crop <- "tropical"
# y.vx <- velox::velox(readr::read_rds(paste0("./data/derived_data/recons/all_",crop,".rds"))[[1]])
# y.vx$extract(japan_thai,
#              fun = function(i){sum(!is.na(i))})[,1]

japan_thai_data <-
  c("Tropical Japonica" = "tropical_japonica",
    "Temperate Japonica" = "temperate_japonica",
    "Indica" = "indica") %>%
  purrr::map(
    function(crop){
      readr::read_rds(paste0(derived_data,"recons/all_",crop,".rds")) %>%
        purrr::map(
          function(y){
          y.vx <- velox::velox(y)
          y.vx$extract(japan_thai,
                       fun = function(i){mean(i, na.rm = TRUE)}) %>%
            t() %>%
            tibble::as_tibble() %>%
            magrittr::set_names(japan_thai$NAME) %>%
            dplyr::mutate(`Years BP` = 
                            names(y) %>% 
                            stringr::str_remove("X") %>% 
                            as.integer()) %>%
            dplyr::select(`Years BP`,
                          dplyr::everything())
        }) 
    }) %>%
  purrr::map(dplyr::bind_rows,
             .id = "Series") %>%
  dplyr::bind_rows(.id = "Crop") %>%
  tidyr::gather(Country, `Probability of being in niche`,-Crop:-`Years BP`) %>%
  tidyr::spread(Series, `Probability of being in niche`)

japan_thai_data %>%
  ggplot2::ggplot(ggplot2::aes(x = `Years BP`,
                               color = Country)) +
  ggplot2::geom_ribbon(ggplot2::aes(ymin = Z_Lower,
                                    ymax = Z_Upper,
                                    fill = Country),
                       color = NA,
                       alpha = 0.3) +
  ggplot2::geom_line(ggplot2::aes(y = Z)) +
  ggplot2::scale_x_reverse(limits = c(5510,1010),
                           breaks = seq(5510,1010,-500),
                           expand = c(0,0)) +
  ggplot2::scale_y_continuous(name = "Probability of being in niche",
                              limits = c(0,100),
                              expand = c(0,0)) +
  
  ggplot2::facet_wrap(facets = "Crop",
                      ncol = 1) +
  ggplot2::theme_minimal() +
  ggplot2::theme(legend.position="bottom",
                 plot.margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0, unit = "pt"))

ggplot2::ggsave(paste0(figures,"supplementary_figure_20b.pdf"),
                width = 6.5,
                height = 6,
                device = cairo_pdf)


smooth.preds <- function(y){
  y[y<0] <- 0
  y[y>1] <- 1
  y <- loess(isoreg(y~sample.points)$yf~sample.points, span=0.1, degree=1) %>%
    skinny.loess()
  return(y)
}

GHCN.data.final$spatial %>%
  sf::st_as_sf() %>%
  sf::st_transform(sf::st_crs(japan_thai)) %>%
  sf::st_intersection(japan_thai) %>%
  dplyr::select(ID) %>%
  tibble::as_tibble() %>%
  sf::st_as_sf() %>%
  dplyr::left_join(
    GHCN.GDD.incremented.sd[[1]] %>%
      dplyr::group_by(ID) %>%
      dplyr::summarise(`GDD Model` = list(loess(isoreg(GDD ~ SD_change)$yf ~ sample.points, span=0.1, degree=1)))
  ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(`GDD` = list(
    marcott2013 %>%
      dplyr::select(`Years BP` = YearBP,
                    Z_Lower,
                    Z,
                    Z_Upper) %>%
      dplyr::mutate_at(.vars = vars(Z_Lower,
                                    Z,
                                    Z_Upper),
                       .funs = function(x){
                         predict(`GDD Model`,newdata = x)
                       })
  )) %>%
  sf::st_as_sf() %>%
  dplyr::select(GDD) %>%
  dplyr::ungroup() %>%
  sf::st_intersection(japan_thai) %>%
  dplyr::rename(Country = NAME) %>%
  sf::st_drop_geometry() %>%
  tidyr::unnest(cols = c(GDD)) %>%
  dplyr::group_by(Country,`Years BP`) %>%
  dplyr::summarise_all(mean, na.rm = TRUE) %>%
  ggplot2::ggplot(ggplot2::aes(x = `Years BP`,
                               color = Country)) +
  ggplot2::geom_ribbon(ggplot2::aes(ymin = Z_Lower,
                                    ymax = Z_Upper,
                                    fill = Country),
                       color = NA,
                       alpha = 0.3) +
  ggplot2::geom_line(ggplot2::aes(y = Z)) +
  ggplot2::scale_x_reverse(limits = c(5510, 1010),
                           breaks = seq(5510, 1010, -500),
                           expand = c(0,0)) +
  ggplot2::scale_y_continuous(name = "Growing Degree Days",
                              limits = c(0,8000),
                              expand = c(0,0)) +
  ggplot2::theme_minimal() +
  ggplot2::theme(legend.position="bottom",
                 plot.margin = ggplot2::margin(t = 10, r = 10, b = 0, l = 0, unit = "pt"))

ggplot2::ggsave(paste0(figures,"supplementary_figure_20a.pdf"),
                width = 6.5,
                height = 6.5/1.618,
                device = cairo_pdf)

```

## Export figure source data for Nature Plants
```{r export graph data}
# Force Raster to load large rasters into memory
raster::rasterOptions(chunksize = 2e+09,
                      maxmemory = 2e+10)

dir.create(paste0(figures,"source_data/"))

crop_GDD %$%
  cultivar %>%
  purrr::map(function(cultivar){
    rasts <- 
      readr::read_rds(paste0(derived_data,"recons/",cultivar,"_recons.rds")) %>%
      purrr::map_dfr(function(x){
        # the_brick <- 
        # x %>%
        # magrittr::extract2(which(.@z$`Years BP` > 1000))
        
        mean.temporal <- 
          raster::cellStats(x, mean, na.rm = T) %>%
          round(digits = 2)
        
        ci.temporal <- 
          raster::cellStats(x,
                            stats::quantile,
                            probs = c(0.25, 0.5, 0.75),
                            na.rm = T)
        
        tibble::tibble(`Years BP` = x@z$`Years BP`,
                       `Spatial Mean` = mean.temporal,
                       `Spatial Lower IQR` = ci.temporal[1,],
                       `Spatial Median` = ci.temporal[2,],
                       `Spatial Upper IQR` = ci.temporal[3,])
        
      },
      .id = "Marcott et al. 2013 Reconstruction") %>%
      dplyr::mutate(`Marcott et al. 2013 Reconstruction`  = `Marcott et al. 2013 Reconstruction` %>%
                      factor(levels = c("Z","Z_Lower","Z_Upper"),
                             ordered = TRUE) %>%
                      forcats::fct_recode("Mean" = "Z",
                                          "1σ Lower" = "Z_Lower",
                                          "1σ Upper" = "Z_Upper")) %>%
      dplyr::select(`Years BP`, dplyr::everything())
    
  }) %>%
  magrittr::set_names(crop_GDD$cultivar_long) %>%
  magrittr::extract(.,sort(names(.))) %>%
  writexl::write_xlsx(paste0(figures,"source_data/figure_3e_cultivar_graph_data.xlsx"))

## Export raster data for Figure 3
readr::read_rds(paste0(derived_data,"recons/tropical_japonica_120_recons.rds")) %$%
  Z %>%
  magrittr::extract2(paste0("X", c(4410, 3550))) %>%
  raster::unstack() %>%
  magrittr::set_names(.,c(4410, 3550)) %>%
  purrr::iwalk(~raster::writeRaster(.x, paste0(figures,"source_data/figure_3_raster_data_",.y,".asc")))

```

## Colophon {-}

This report was generated on `r Sys.time()` using the following computational environment and dependencies: 

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r git details, cache = FALSE}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("..")
```
